---
title: "Introduction to R"
author: "Subrata Paul"
subtitle: "Very Basic"
description: "Important basics of R"
footer: "[🔗 spaul-genetics.github.io/RStat4Arts](https://spaul-genetics.github.io/RStat4Arts)"
format: 
  revealjs:
    theme: solarized
    transition: fade
    slide-number: true
    incremental: false 
    chalkboard: true
    center: true
execute:
  freeze: auto
  echo: true
---


# Welcome

```{r}
#| echo: false
#| message: false
#| warning: false
library(tidyverse)
```

## R

::: incremental
-   Programming Langulage
-   Other langulages: html, java, python
-   Purpose: Data and statistical analysis
-   What can be done: almost anything
-   Limitation: Memory usage
:::

## Getting Started

* Install R from : <https://www.r-project.org>

* Install Rstudio from : <https://posit.co/download/rstudio-desktop/>

* Also install Quarto: <https://quarto.org/docs/get-started/>

# Practical Intro to Rstudio

## Learned

::: incremental
-   How to launch Rstudio
-   Components of Rstudio
-   Where to write code and where to execute it
-   File navigation
:::

# R Object

## Assignment 

* Use either `<-` or `=`

```{r}
x <- 3
print(x)
```

```{r}
y = 5
y
```


## Vector

```{r}
x <- c(1, 4, 10, 11)
x
```

- Element wise operation

```{r}
x * 2
x + 3
x - 3
y = 3*x + 5
print(y)
```

## Class

```{r}
class(1)
class('a')
class(TRUE)
```

## Special Class 1

```{r}
x = 4
class(x)
x = as.integer(x)
class(x)
```

## Special Class 2

```{r}
gender = sample(c('Male', 'Female'), size = 20, replace = T)
class(gender)
gender = as.factor(gender)
# Use ?as.factor to get details 
class(gender)
```



## Slicing

```{r}
LETTERS
LETTERS[2:5]
LETTERS[c(1,5,10)]
LETTERS[c(1,4,6:9)]
```

## Subset

```{r}
Zila = sample(c('Munshiganj', 'Dhaka','Gazipur'), size = 10, replace = T)
Zila
Income = sample(40000:100000, size = 10, replace = T)
Income

Zila == 'Munshiganj'
Income[Zila == 'Munshiganj']
```


## Getting Help

* ?command
  - ?sum
  - ?for
  
* ??something

## Using Function

```{r}
pnorm(0.5)
pnorm(0.5, 0, 1)
pnorm(0.5, mean = 0, sd = 1)
pnorm(0.5, 1, 0)
```


## Sequence

```{r}
1:5
seq(1, 5)
seq(1,5,2)
seq(from = 1, to = 5, length.out = 4)
```

# Data Frame

## Reading

```{r}
mydata = read.csv('~/Documents/All_R_projects/RStat4Arts/data/toy.example1.csv')
head(mydata)
```

## Excel File

```{r eval=F}
install.packages('readxl')
```

```{r}
library(readxl)
some_other_data = read_excel('../data/toy.example1.xlsx')
head(some_other_data)
```

## Without Loading Package

```{r}
some.new.data <- readxl::read_excel('../data/toy.example1.xlsx')
```

## Access variable

```{r}
mydata$Zila
```

## Subset Data 1

```{r}
mydata[1:5,1:3]
```

## Subset Data 2

```{r}
mydata[1:5,]
```

## Subset Data 3

```{r}
mydata[c(1,3,6,7),c('Person', 'Zila')]
```

## Subset Data 4

```{r}
mydata[mydata$Zila=='Munshiganj',]
```

## Subset Data 5

```{r}
mydata[mydata$Income>38000,]
```

## Subset Data 6

```{r}
interesting_persons = c("Person1", 'Person10', 'Person21', 'Person35')
mydata[mydata$Person%in%interesting_persons,]
```

## Add Variable

```{r}
mydata$Family.Size = sample(1:10, nrow(mydata), replace = T)
head(mydata)
```

## Add Variable 

```{r}
mydata$Per.Head.Income = mydata$Income/mydata$Family.Size
```

# Basic Stats

## Central Tendency


```{r}
plot(iris$Sepal.Length)
abline(h = seq(4.5, 8, 0.5))
```

```{r}
table(cut(iris$Sepal.Length, seq(4,8,0.5)))
```

## Central Tendency (Cont.)

```{r}
hist(iris$Sepal.Length, breaks = 20)
```

## Measures of Center

::: incremental

* Mean
* Median
* Mode

:::

## Spread

```{r echo=F}
library(ggplot2)
plot_data = data.frame(Run = c(round(rnorm(50, mean = 50, sd = 5)),round(rnorm(50, mean = 50, sd = 40))), Player = c('Messi', 'Hassi'))
plot_data$Run[plot_data$Run<0]<-0
ggplot(data = plot_data, aes(y = Run, x = Player))+geom_jitter(width = 0.1)
```

## Measures of spread

* Range
* Standard Deviation
* Variance

## Boxplot

```{r}
ggplot(data = plot_data, aes(y = Run, x = Player, fill = Player))+geom_boxplot()+ggthemes::scale_fill_tableau()
```


# Plotting

## R Default

```{r}
plot(iris$Sepal.Length, iris$Sepal.Width)
```

## ggplot2

```{r eval=F}
install.packages('ggplot2')
```

```{r}
library(ggplot2)

ggplot(data = iris)+
  aes(x = Sepal.Length, y = Sepal.Width)+
  geom_point()
```

## Color points

```{r}
ggplot(data = iris)+
  aes(x = Sepal.Length, y = Sepal.Width, color = Species)+
  geom_point()
```

## Theme

```{r}
ggplot(data = iris)+
  aes(x = Sepal.Length, y = Sepal.Width, color = Species)+
  geom_point()+
  theme_bw()+
  labs(x = 'Sepal Length', y = 'Sepal Width')+
  theme(legend.position = 'top', 
        axis.title = element_text(face = 'bold'))
```
